---
title: "Ukraine Appropriations Analysis"
author: "Justin S."
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: TRUE
    toc_depth: 4
    highlight: tango
    code_folding: hide
    lightbox: TRUE
    gallery: TRUE
    toc_float: # a float toc will stick to the sidebar when scrolling
      collapsed: false

---

On x, Congress approved its omnibus spending bill, which included $13.6 billion in aid for Ukraine. This brief analysis looks at a breakdown of the [Ukraine approprations](https://appropriations.house.gov/sites/democrats.appropriations.house.gov/files/Ukraine%20Supplemental%20Summary.pdf), looking at the line items in the bill.

```{r fig.align='center', fig.height=6, fig.width=6, message=FALSE, warning=FALSE, include=FALSE}
# Set our plot specifications for the rest of the document.
knitr::opts_chunk$set(fig.width = 9,
                      fig.height = 6,
                      fig.align = "center",
                      # Set our code specifications for the rest of the document
                      echo = F,
                      warning = F,
                      message = F)
```


```{r set-up_and_clean_data, include=FALSE}
###########################################################################
## Set Up -----------------------------------------------------------------
###########################################################################
# Bring in packages
suppressMessages(library("pacman"))
pacman::p_load("tidyverse", # Used for data wrangling,
               "tidyr", # Used for data cleaning,
               DataExplorer, # Exploratory data analysis
               readr,
               "ggplot2", # Used for visualizations,
               "here", # Used for navigating project structure
               "readxl", # Used for loading excel files,
               "pander", # Used for pretty tables,
               "kableExtra", # Used for RMarkdown formatting
               "janitor", # Used for data cleaning,
               "stringr")  # Used for string manipulation

# Get rid of scientific notation
options(scipen = 999)

```


Let's start by taking a look at the data.

```{r display_data}
########################################################################
### View Data --------------------------------------------------------
########################################################################
ukraine <- read_excel(here::here('Data/Ukraine Appropriations Breakdown.xlsx'))

# Let's view the data output
# In any kable outputs, display NAs as blanks
opts <- options(knitr.kable.NA = "")

ukraine %>% 
  head(25) %>%
  # Fix up the headers by replacing the underscores with spaces
  rename_all(funs(str_replace_all(., "_", " "))) %>% 
  # Make everything proper capitalization
  # rename_all(funs(str_to_title)) %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = F,
                font_size = 10) %>%
  # Make the header row bold and black so it's easier to read
  row_spec(0, bold = T, color = "black") %>% 
  scroll_box(height = "400px", width = "100%")
```




```{r clean_data}
ukraine_cleaned <- ukraine %>% 
  janitor::clean_names() %>% 
  # Convert the appropriations column into a numeric
  mutate(appropriations_cleaned = stringr::str_remove(appropriations, '[$]'),
         appropriations_cleaned = if_else(
           stringr::str_detect(appropriations_cleaned, 'million'),
           readr::parse_number(appropriations) * 1000000,
           readr::parse_number(appropriations) * 1000000000
         ))
  
```


```{r}
ukraine_cleaned %>%
  # Group by title
  group_by(title) %>%
  # Count everything up!
  summarise(total_appr = sum(appropriations_cleaned)/1000000) %>%
  ungroup() %>%
  # Start our visualization, creating our groups by party affiliation
  ggplot(aes(x = reorder(title, -total_appr), y = total_appr, fill = total_appr)) +
  geom_col() +
  # Change the theme to classic
  theme_classic() +
  scale_fill_gradient(low = 'gray', high = 'slateblue', name = 'Appropriations') +
  # Let's change the names of the axes and title
    xlab("Title of Bill") +
    ylab("Total Appropriations (millions of $)") +
    labs(title = "Total Appropriations in Congressional Spending Bill (H.R. 2471) by Title",
         subtitle = paste0("Allocations are broken out across ", n_distinct(ukraine_cleaned$title), " Titles and ", nrow(ukraine_cleaned), " line items."),
         caption = "Data is gathered from the House of Representatives Appropriations page at https://appropriations.house.gov/news/press-releases/house-passes-government-funding-legislation") +
    # format our title and subtitle
    theme(plot.title = element_text(hjust = 0, color = "slateblue4"),
          plot.subtitle = element_text(hjust = 0, color = "slateblue2", size = 10),
          plot.caption = element_text(color = "dark gray", size = 10, face = "italic"))
```

Here we can see that Titles III and Titles VI overwhelm the level of appopriations doled out by Congress. These include:

```{r}
allocations_by_title <- ukraine_cleaned %>% 
  group_by(title, title_description) %>% 
  summarise('Total Appropriations ($)' = sum(appropriations_cleaned)) %>% 
  ungroup()

allocations_by_title %>%
  select(title, title_description, 'Total Appropriations ($)') %>% 
  filter((title == 'Title III') | (title == 'Title VI')) %>% 
  pander()
```

From here on out, because defense and state operations constitute the overwhelming majority of 
appropriations, I'll visualize them separately. Here's what the same bar chart looks like when separated:

```{r}
ukraine_grouped <- ukraine_cleaned %>% 
  mutate(title_grouped = case_when(
    title == 'Title III' ~ 'Defense/State',
    title == 'Title VI' ~ 'Defense/State',
    TRUE ~ 'Other'
  ))

ukraine_grouped %>%
  # Group by title
  group_by(title, title_grouped) %>%
  # Count everything up!
  summarise(total_appr = sum(appropriations_cleaned)/1000000) %>%
  ungroup() %>%
  # Start our visualization, creating our groups by party affiliation
  ggplot(aes(x = reorder(title, -total_appr), y = total_appr, fill = total_appr)) +
  geom_col() +
  # Change the theme to classic
  theme_classic() +
  facet_wrap(~title_grouped, scales = 'free') + 
  scale_fill_gradient(low = 'gray', high = 'slateblue', name = 'Appropriations') +
  # Let's change the names of the axes and title
    xlab("Title of Bill") +
    ylab("Total Appropriations (millions of $)") +
    labs(title = "Total Appropriations in Congressional Spending Bill (H.R. 2471) by Title",
         subtitle = paste0("Allocations are broken out across ", n_distinct(ukraine_cleaned$title), " Titles and ", nrow(ukraine_cleaned), " line items."),
         caption = "Data is gathered from the House of Representatives Appropriations page at https://appropriations.house.gov/news/press-releases/house-passes-government-funding-legislation") +
    # format our title and subtitle
    theme(plot.title = element_text(hjust = 0, color = "slateblue4"),
          plot.subtitle = element_text(hjust = 0, color = "slateblue2", size = 10),
          plot.caption = element_text(color = "dark gray", size = 10, face = "italic"))
```

Within the Defense and State portions, here are some of the most notable line-items for appropriations:

```{r ds_line_items}
ukraine_grouped %>% 
  filter(title_grouped == 'Defense/State') %>% 
  arrange(desc(appropriations_cleaned)) %>% 
  select(title_description, appropriations, description) %>% 
  head() %>% 
  pander()
```
